---
title: "Capstone: Logistic Regression"
author: "Andrew Knittle"
date: "`r paste('Generated on', Sys.Date())`" 
output: 
  html_document: 
    # try readable or cosmo
    # default (default), try kable, tibble, or paged as an option
    # Options are none (default), show, and hide
    # default is null, try my_style_file.css
    toc: TRUE
    toc_depth: 3
    toc_float: FALSE
    highlight: haddock
    theme: flatly
    df_print: paged
    code_folding: none
    self_contained: TRUE
---

```{r setup, include=FALSE}
# --------------------------------------
# Clean and Prep Environment
rm(list=ls())
set.seed(123)

# --------------------------------------
# Load Libraries
library(pacman)
p_load(ggplot2)
p_load(ISLR)
p_load(dplyr)
p_load(psych)
p_load(ggpubr)
p_load(kableExtra)
p_load(corrplot)
p_load(tidyverse)
p_load(caret)
p_load(MASS)
p_load(glm.predict)
p_load(glmnet)
p_load(boot)
p_load(pROC)
# Formatting----------------------------
p_load(kable)
p_load(kableExtra)

# --------------------------------------
# Read Data
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

dustData = as.data.frame(read.csv('Data/dustData_Reduced.csv',header=T,sep=',')) # OG dataset
dustData <- dustData[-1]
dustData$CASESTAT <- as.factor(dustData$CASESTAT)
dustData <- na.omit(dustData[,c(1:27, 30)])

logDustData = as.data.frame(read.csv('Data/logDustData_Reduced.csv',header=T,sep=','))
logDustData <- logDustData[-1]
logDustData <- logDustData[-1]
logDustData$CASESTAT <- as.factor(logDustData$CASESTAT)
logDustData <- na.omit(logDustData[,c(1:27, 30)])

train_DustData <- as.data.frame(read.csv('Data/Training Set_0.9_May 25 2021.csv',header=T,sep=','))[-1]
#train_DustData$CASESTAT <- as.numeric(train_DustData$CASESTAT)
train_DustData$CASESTAT <- as.factor(train_DustData$CASESTAT)
train_DustData <- na.omit(train_DustData[,c(1:27, 30)])

# --------------------------------------
# Read helper script
source("Exploratory_Support_Scripts.R")
source("CrossValidation_Library.R")
source("Script_Library.R")



```

# Exploring the Data

```{r Correlation Matrix, echo=FALSE, fig.cap= "Correlation Matrix", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}


# Correlation matrix
# Probably looking at one of the Benz ones
#cor(dustData_Reduced[1:26], use="complete.obs")
corrplot(cor(dustData[1:26], use="complete.obs"))
# corrplot(M, method="circle")
# http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram


```


As we can see there are many variables that are highly correlated. The goal now is to find what variables we need and which we can throw out. We will break the variables into groups reflecting the output from the correlation graph.

# Finding the Right Variables

```{r Group 1, echo=FALSE, fig.cap= "Group 1", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE, results='asis'}

# Group 1:
# BENZ_A_ANTHR, BENZ_B_FLUOR, BENZ_K_FLUOR, BENZ_A_PYRENE, CHRYSENE, DIBENZ_ANTHR, IDENO_PYR
BENZ_A_ANTHR <- glm(CASESTAT ~ BENZ_A_ANTHR_1, data = train_DustData, family = "binomial")
BENZ_B_FLUOR <- glm(CASESTAT ~ BENZ_B_FLUOR_1, data = train_DustData, family = "binomial")
BENZ_K_FLUOR <- glm(CASESTAT ~ BENZ_K_FLUOR_1, data = train_DustData, family = "binomial")
BENZ_A_PYRENE <- glm(CASESTAT ~ BENZ_A_PYRENE_1, data = train_DustData, family = "binomial")

CHRYSENE <- glm(CASESTAT ~ CHRYSENE_1, data = train_DustData, family = "binomial")
DIBENZ_ANTHR <- glm(CASESTAT ~ DIBENZ_ANTHR_1, data = train_DustData, family = "binomial")
IDENO_PYR <- glm(CASESTAT ~ IDENO_PYR_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(BENZ_A_ANTHR)
cat("---------------------------------\n")
summary(BENZ_B_FLUOR)
cat("---------------------------------\n")
summary(BENZ_K_FLUOR)
cat("---------------------------------\n")
summary(BENZ_A_PYRENE)
cat("---------------------------------\n")


summary(CHRYSENE)
cat("---------------------------------\n")
summary(DIBENZ_ANTHR)
cat("---------------------------------\n")
summary(IDENO_PYR)

Group1_Chems <- c("BENZ_A_ANTHR", "BENZ_B_FLUOR", "BENZ_K_FLUOR", "BENZ_A_PYRENE", "CHRYSENE", "DIBENZ_ANTHR", "IDENO_PYR")
# AIC <- c(extractAIC(BENZ_A_ANTHR), extractAIC(BENZ_B_FLUOR), extractAIC(BENZ_K_FLUOR), extractAIC(BENZ_A_PYRENE), extractAIC(CHRYSENE), 
#          extractAIC(DIBENZ_ANTHR), extractAIC(IDENO_PYR))
# cbind.data.frame(Group1_Chems, AIC)
```

Group 1 Winner: CHRYSENE  Estimate -0.07203, P-Value: 0.519, AIC: 1464.7


```{r Group 2, echo=FALSE, fig.cap= "Group 2", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}

# ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
# Group 2:
# PCB105, PCB138, PCB153
PCB105 <- glm(CASESTAT ~ PCB105_1, data = train_DustData, family = "binomial")
PCB138 <- glm(CASESTAT ~ PCB138_1, data = train_DustData, family = "binomial")
PCB153 <- glm(CASESTAT ~ PCB153_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(PCB105)
cat("---------------------------------\n")
summary(PCB138)
cat("---------------------------------\n")
summary(PCB153)


Group2_Chems <- c("PCB105", "PCB138", "PCB153")
```

Group 2 Winner: PCB105: Estimate: 0.1194, P-Value: 0.2650, AIC: 1463.8


```{r Group 3, echo=FALSE, fig.cap= "Group 3", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}
# ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
# Group 3:
# PCB170, PCB180 <May need to look at other PCB's as well>
PCB170 <- glm(CASESTAT ~ PCB170_1, data = train_DustData, family = "binomial")
PCB180 <- glm(CASESTAT ~ PCB180_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(PCB170)
cat("---------------------------------\n")
summary(PCB180)

Group3_Chems <- c("PCB170", "PCB180")
```

Group 3 Winner: PCB180: Estimate: 0.18404, P-Value: 0.082, AIC: 1462


```{r Group 4, echo=FALSE, fig.cap= "Group 4", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}
# ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
# Group 4:
# A_CHLORDANE, G_CHLORDANE

A_CHLORDANE <- glm(CASESTAT ~ A_CHLORDANE_1, data = train_DustData, family = "binomial")
G_CHLORDANE <- glm(CASESTAT ~ G_CHLORDANE_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(A_CHLORDANE)
cat("---------------------------------\n")
summary(G_CHLORDANE)

Group4_Chems <- c("A_CHLORDANE", "G_CHLORDANE")
```

Group 4 Winner: A_CHLORDANE: Estimate: 0.09623, P-Value: 0.253, AIC: 1463.8


```{r Group 5, echo=FALSE, fig.cap= "Group 5", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}
# ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
# Group 5:
# DDE, DDT

DDE <- glm(CASESTAT ~ DDE_1, data = train_DustData, family = "binomial")
DDT <- glm(CASESTAT ~ DDT_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(DDE)
cat("---------------------------------\n")
summary(DDT)

Group5_Chems <- c("DDE", "DDT")
```

Group 5 Winner: DDE: Estimate: 0.21830, P-Value: 0.0661, AIC: 1461.7


```{r Group 6, echo=FALSE, fig.cap= "Group 6", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}
# ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
# Group 6:
# CISPERMETHRIN, TRANSPERMETHRIN

CISPERMETHRIN <- glm(CASESTAT ~ CISPERMETHRIN_1, data = train_DustData, family = "binomial")
TRANSPERMETHRIN <- glm(CASESTAT ~ TRANSPERMETHRIN_1, data = train_DustData, family = "binomial")

# -------------------------------------------
cat("---------------------------------\n")
summary(CISPERMETHRIN)
cat("---------------------------------\n")
summary(TRANSPERMETHRIN)

Group6_Chems <- c("CISPERMETHRIN", "TRANSPERMETHRIN")


```

Group 6 Winner: TRANSPERMETHRIN: Estimate: -0.02564, P-Value: 0.618, AIC: 1464.8


# Building Models

## Stepwise Selection

This model will be built using stepwise selection, using AIC as a criteria.

```{r Stepwise Model, echo=FALSE, fig.cap= "Stepwise Model", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}



stepwiseModel1 <- glm(CASESTAT ~ ., data = train_DustData, family = "binomial") %>% stepAIC(trace = FALSE)
cat("---------------------------------\n")
summary(stepwiseModel1)

```


## Using Hand Picked Variables

Using the variables selected from before this is the current model

```{r Building Model, echo=FALSE, fig.cap= "Building Model", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}

model1 <- glm(CASESTAT ~ CHRYSENE_1 + PCB105_1 + PCB180_1 + A_CHLORDANE_1 + DDE_1 + TRANSPERMETHRIN_1, data = train_DustData, family = "binomial")
cat("---------------------------------\n")
summary(model1)

```



# Comparing Models


## Handselected Model

```{r Comparing Handselected Model Training, echo=FALSE, fig.cap= "Comparing Handselected Models Training", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}



roughPerformanceTest(train_DustData, "CASESTAT", model1)


# ----------------------------------------

confusion.result <- confusionBuilder(train_DustData$CASESTAT, model1$fitted.values, 0.5)
confusion.result


tibble(Model=c("Logistic Model"), Error_Rate = c(error.rater(train_DustData, model1, 10)))
```


## Stepwise Model


```{r Comparing Stepwise Model Training, echo=FALSE, fig.cap= "Comparing Stepwise Models Training", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}

# ----------------------------------------

# roughPerformanceTest(train_DustData, "CASESTAT", stepwiseModel1)

confusion.result <- confusionBuilder(train_DustData$CASESTAT, stepwiseModel1$fitted.values, 0.5)
confusion.result

tibble(Model=c("Logistic Model"), Error_Rate = c(error.rater(train_DustData, stepwiseModel1, 10)))
# TODO: Out of curiosity throw everything into a logistic model

# rocFull <- rocBuilder(train_DustData$CASESTAT, stepwiseModel1$fitted.values, toleranceVec = seq(0,1,0.001))
# rocFull[[2]]
# rocDF <- rocFull[[1]]
# simpleAUC(rocDF$`True Positive Rate`, rocDF$`False Positive Rate`)
# 
# 
# rocByLibrary(stepwiseModel1, train_DustData, train_DustData$CASESTAT, type="response")

```



# Conclusion

The Stepwise Model performed best compare to "By Hand" model in terms of error rate for the test set using 10-Fold Cross Validation.

```{r Final Model, echo=FALSE, fig.cap= "Final Model", out.height="100%", out.width="100%", warning=FALSE, comment=FALSE, warning=FALSE,  results='asis'}


# summary(stepwiseModel1.test)
# 
# # ----------------------------------------
# 
# 
# confusion.result
# errorRateTable
```

CASESTAT ~ 1.4191 + (0.1780)D24 + (0.5941)BENZ_K_FLUOR + (-3.3352)BENZ_A_PYRENE + (-2.9114)CHRYSENE + 
           (5.2359)IDENO_PYR + (-0.1204)PCB180 + (0.9109)DDE + (-0.2635)DDT + (-0.1117)DIAZINON + (-0.1794)METHOXYCHLOR

We see that only 3 variables are statistically significant when using data from the testset; BENZ_A_PYRENE, CHRYSENE, and IDENO_PYR. Curiously only one of the chemicals has a positive coefficient estimate in log odds.

## Method

Logistic Regression was chosen because of the binary dependent variable and the simplicity of implementing it. Logistic Regression gives us the probability that someone will likely contract cancer when introduced to the chemicals in the model, and based on that probability we can classify a person developing cancer or not.

Each coefficient represents the log odds of one unit of chemicals making a subject have cancer. So for example from the model that was produced earlier the log odds of developing cancer increases by 5.23 for every one unit of IDENO_PYR.













